### 2023-10-11 main.tex
P assorbimento

	\langle n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}n_{s_{3}}^{y_{3}}\rangle_{stat}^{c}&=\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}\right]-\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{3}}^{y_{3}}\right]-\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{3}}^{y_{3}}\right]\mathbb{E}\left[n_{s_{2}}^{y_{2}}\right]\\&-\mathbb{E}\left[n_{s_{3}}^{y_{3}}n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{1}}^{y_{1}}\right]+2\mathbb{E}\left[n_{s_{1}}^{y_{1}}\right]\mathbb{E}\left[n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{3}}^{y_{3}}\right]

@@

\textbf{Remark}: the first and second cumulants, computed in \eqref{one-pts-corr} and \eqref{two-pts-corr} respectively, match with the ones found in Subsection 4.3 of \cite{vanicat2017exact}.

\section{Extensions for the non-integrable chain}
In Sections \ref{sectionIntegrabiliy} we have shown that, on the geometry of a chain and with hard-core interaction the model is exactly solvable using to integrability.  
The aim of this section is studying some property of the process when integrability is lost. We fix $\nu\geq 1$ and, for the sake of simplicity, we keep the geometry of a chain. We provide the following extensions: 
\begin{enumerate}
	%\item \textit{Arbitrary correlations}: exploiting duality, the steady state correlations can be expressed in terms of absorption probabilities of the dual particles. These probability are hard to find in this partial exclusion situation, however correlation can be formally written. Subsection \ref{Subsection-ss-nonI}.
	\item \textit{Instantaneous thermalization}: starting from the multi-species stirring dynamics, we define the thermalized process, where particle are redistribute in each bond with a measure that is the reversible one conditioned to the conservation of the number of particles on this bond. Also for this thermalized process we prove duality. Subsection \ref{Subsection-Thermalized}.
	\item \textit{Reaction-diffusion process}: a further dynamics is added. Particle can now change type. The model is carefully constructed with the goal of having the average densities evolving as a system of $N-1$ coupled difference-differential equation, providing a generalization of the model studied in \cite{casini2022uphill}. Duality for the boundary driven process is also proved. Subsection \ref{Subsection-RD}
\end{enumerate}

@@

	\paragraph{Proof of formula \eqref{steadyStateH}} As a consequence of \eqref{S1-Inverse}, we only need to show how the transformation $S_{1}^{-1}$ acts on the vector $|\Psi^{'}\rangle$. Here is more convenient to consider the matrix $\mathcal{S}_{1}^{-1}$ defined in \eqref{S2Gen}, acting on a single site, and make explicitly the tensor product over the chain. This matrix $\mathcal{S}_{1}$ is given by
	\begin{equation}\label{S1-L1}
		\mathcal{S}_{1}^{-1}=\mathbbm{1}+(-1)\sum_{a=1}^{N-1}e_{Na}
	\end{equation}
thus
\begin{equation}
\mathcal{S}_{1}^{-1}|\mu\rangle=|\mu\rangle+(-1)\delta_{\mu<N}|N\rangle
\end{equation}
By taking the tensor product over the chain we obtain 
\begin{equation}
	\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\left(|\mu_{1}\rangle+(-1)\delta_{\mu_{1}<N}|N\rangle\right)\otimes \ldots \otimes \left(|\mu_{L}\rangle+(-1)\delta_{\mu_{L}<N}|N\rangle\right)
\end{equation}
therefore, projecting on $\langle \bm{\tau}|$ we have
\begin{equation}
	\langle \bm{\tau}|	\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\prod_{x=1}^{L}\left(\delta_{\mu_{x},\tau_{x}}+(-1)\delta_{\tau_{x},N}\sum_{i_{x}=1}^{N-1}\delta_{\mu_{x},i_{x}}\right)
\end{equation}
\begin{comment}
	Using the notation introduced in \eqref{notation1} and \eqref{notation2} we have that 
	\begin{equation}
		\begin{split}
			\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}&=\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L}}^{\tau_{1},\ldots,\tau_{L}}\right.
			\\&
			+\left.\sum_{b=1}^{L}(-1)^{b}\sum_{q_{1}=1}^{L}\sum_{q_{2}=q_{1}+1}^{L}\ldots\sum_{q_{b}=q_{b-1}+1}^{L}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{r}}=1}^{N-1}e_{\eta_{1},\ldots,\eta_{L}}^{\theta_{1},\ldots,\theta_{L}}\right\} 
		\end{split}
	\end{equation}
	where we have introduced, $\forall x\in \{1,\ldots,L\}$
	\begin{equation}
		%\eta_{x}=\begin{cases}
			%	\tau_{x}\qquad &\text{if}\quad x\neq 1,\ldots,b\\
			%	d_{x}\qquad &\text{otherwise}
			%\end{cases}\qquad
			\theta_{x}(b)=\begin{cases}
				\tau_{x}\qquad &\text{if}\quad x\notin \{q_{1},\ldots,q_{b}\}\\
				N \qquad &\text{otherwise}
			\end{cases}
		\end{equation}
		The dependence of $\theta(b)$ with respect to $b$ is usually understood for the sake of notation. 
		This is proved by induction. Indeed, for $L=1$ it is trivially true, by \eqref{S1-L1}. We assume that it is true for arbitrary $L\in \mathbb{N}$ and we prove that it holds for $L+1$. 
		\begin{align*}
			\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}\otimes \mathcal{S}_{1}^{-1}&=
			\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L}}^{\tau_{1},\ldots,\tau_{L}}\right. \\&+\left. \sum_{b=1}^{L}(-1)^{b}\sum_{q_{1}=1}^{L}\sum_{q_{2}=q_{1}+1}^{L}\ldots\sum_{q_{b}=q_{b-1}+1}^{L}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{r}}=1}^{N-1}e_{\theta_{1},\ldots,\theta_{L}}^{\eta_{1},\ldots,\eta_{L}}\right\}
			\\&\otimes
			\left\{ \sum_{\tau_{1}=1}^{N}e_{\tau_{L+1}}^{\tau_{L+1}}+(-1)\delta_{\tau_{L+1},N}\sum_{d_{L+1}=1}^{N-1}e_{1}^{d_{L+1}}\right\}
			\\&=
			\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L+1}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L+1}}^{\tau_{1},\ldots,\tau_{L+1}}\right. \\&+\left. \sum_{b=1}^{L+1}(-1)^{b}\sum_{q_{1}=1}^{L+1}\sum_{q_{2}=q_{1}+1}^{L+1}\ldots\sum_{q_{b}=q_{b-1}+1}^{L+1}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{b}}=1}^{N-1}e_{\theta_{1},\ldots,\theta_{L+1}}^{\eta_{1},\ldots,\eta_{L+1}}\right\}
		\end{align*}
		\begin{comment}

>>>>

	\langle n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}n_{s_{3}}^{y_{3}}\rangle_{stat}^{c}&=\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}\right]-\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{3}}^{y_{3}}\right]-\mathbb{E}\left[n_{s_{1}}^{y_{1}}n_{s_{3}}^{y_{3}}\right]\mathbb{E}\left[n_{s_{2}}^{y_{2}}\right]\nonumber\\&-\mathbb{E}\left[n_{s_{3}}^{y_{3}}n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{1}}^{y_{1}}\right]+2\mathbb{E}\left[n_{s_{1}}^{y_{1}}\right]\mathbb{E}\left[n_{s_{2}}^{y_{2}}\right]\mathbb{E}\left[n_{s_{3}}^{y_{3}}\right]\nonumber

@@

\textbf{Remark}: the first and second cumulants, computed in \eqref{one-pts-corr} and \eqref{two-pts-corr} respectively, match with the ones found in Subsection 4.3 of \cite{vanicat2017exact}. \\ \\
\textbf{How $|\Psi^{'}\rangle$ is related to steady state non-equilibrium correlations and to absorption probabilities?}\\
We aim to compute correlations between non-empty species in the steady state. We know that for any chain length $L$ we have that the steady state $|\Psi\rangle$ is related to $|\Psi^{'}\rangle$ by
\begin{equation*}
	|\Psi\rangle=\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\Psi^{'}\rangle
\end{equation*}
We aim to compute the correlations between non-empty particles of an arbitrary configuration $\bm{\tau}$.
This is equivalent to consider a shorter chain where only non-empty particles are left, while all the sites occupied by the empty $N$ are removed. 
This chain has length $m=\sum_{x=1}^{L}(1-\delta_{\tau_{x},N})$, where all these $m$ sites are occupied by particles of species for $1$ to $N-1$. To find the desired correlations we compute the steady state of this shorter chain. We denote its configuration chain by $\widehat{\bm{\tau}}$. This configuration is obtained from $\bm{\tau}$ by re-labelling properly the occupations from $1$ to $m$. Then, these $m$ point correlations can be computed by finding $\Psi(\widehat{\bm{\tau}})$.\\ \\ To clarify consider the example $N=3$ and $L=4$ where the empty is the species $3$. Take a configuration  
	\bm{\tau}=(\tau_{1}=2,\tau_{2}=3,\tau_{3}=1,\tau_{4}=3)
What is the correlation between the occupied species? To compute this we can define the new chain of length $m=2$ with the configuration $\widehat{\bm{\tau}}=(\widehat{\tau}_{1}=2,\widehat{\tau}_{2}=1)$ and find the steady state $|\Psi\rangle$ of this shorter chain. \\ \\
To compute the steady state of the shorter chain we can use the result that we have in the appendix A in equation \eqref{Tensor-S1}:
\begin{equation*}
		\langle \bm{\tau}|	\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\prod_{x=1}^{L}\left(\delta_{\mu_{x},\tau_{x}}+(-1)\delta_{\tau_{x},N}\sum_{i_{x}=1}^{N-1}\delta_{\mu_{x},i_{x}}\right)
\end{equation*}
Thus, for the shorter chain we have that 
\begin{equation*}
	\langle \widehat{\bm{\tau}}|\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\prod_{x=1}^{m}\left(\delta_{\mu_{x},\tau_{x}}\right)
\end{equation*}
this implies that 
	\Psi(\widehat{\bm{\tau}})=\Psi^{'}(\widehat{\bm{\tau}})
where here both $|\Psi(\widehat{\bm{\tau}})\rangle$ and $|\Psi^{'}(\widehat{\bm{\tau}})\rangle$ are eigenvector with vanishing eigenvalue of chains of length $m$. 
We have computed this $\Psi^{'}(\widehat{\bm{\tau}})$ in \eqref{exactCorrelations} as
\begin{equation*}
	\Psi^{'}(\widehat{\bm{\tau}})\sum_{c_{1},\ldots,c_{m}=0}^{1}
\end{equation*}
\begin{equation*}
\end{equation*}
\begin{equation*}
\end{equation*}
We also know from \eqref{CorrelationByABS-prob} that 
\begin{equation*}
	\Psi(\widehat{\bm{\tau}})= \sum_{c_{1}=0}^{1}\ldots \sum_{c_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{c_{k}}\beta_{a_{k}}^{1-c_{k}}\mathcal{P}(c_{1},\ldots,c_{m},\hat{\bm{\xi}})
\end{equation*}
therefore, we can read off the absorption probabilities. \\ \\
This argument can be adapted for any $1\geq m\leq L$ to compute $m$-point correlations between non-empty species.\\ \\ 
In general $|\Psi^{'}\rangle$ is the eigenvector with vanishing eigenvalue of $H^{'}$, but I believe it cannot be interpreted directly as a steady state, since $H^{'}$ is not a Markov matrix. However, it gives by the previous argument the exact correlations and as a consequence the absorption probabilities...
\section{Extensions for the non-integrable chain}
In Sections \ref{sectionIntegrabiliy} we have shown that, on the geometry of a chain and with hard-core interaction the model is exactly solvable using to integrability.  
The aim of this section is studying some property of the process when integrability is lost. We fix $\nu\geq 1$ and, for the sake of simplicity, we keep the geometry of a chain. We provide the following extensions: 
	%\item \textit{Arbitrary correlations}: exploiting duality, the steady state correlations can be expressed in terms of absorption probabilities of the dual particles. These probability are hard to find in this partial exclusion situation, however correlation can be formally written. Subsection \ref{Subsection-ss-nonI}.
	\item \textit{Instantaneous thermalization}: starting from the multi-species stirring dynamics, we define the thermalized process, where particle are redistribute in each bond with a measure that is the reversible one conditioned to the conservation of the number of particles on this bond. Also for this thermalized process we prove duality. Subsection \ref{Subsection-Thermalized}.
	\item \textit{Reaction-diffusion process}: a further dynamics is added. Particle can now change type. The model is carefully constructed with the goal of having the average densities evolving as a system of $N-1$ coupled difference-differential equation, providing a generalization of the model studied in \cite{casini2022uphill}. Duality for the boundary driven process is also proved. Subsection \ref{Subsection-RD}

@@

	\paragraph{Proof of formula \eqref{steadyStateH}} As a consequence of \eqref{S1-Inverse}, we only need to show how the transformation $S_{1}^{-1}$ acts on the vector $|\Psi^{'}\rangle$. Here is more convenient to consider the matrix $\mathcal{S}_{1}^{-1}$ defined in \eqref{S2Gen}, acting on a single site, and make explicitly the tensor product over the chain. This matrix $\mathcal{S}_{1}$ is given by
	\begin{equation}\label{S1-L1}
		\mathcal{S}_{1}^{-1}=\mathbbm{1}+(-1)\sum_{a=1}^{N-1}e_{Na}
	\end{equation}
thus
\begin{equation}
\mathcal{S}_{1}^{-1}|\mu\rangle=|\mu\rangle+(-1)\delta_{\mu<N}|N\rangle
\end{equation}
By taking the tensor product over the chain we obtain 
\begin{equation}
	\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\left(|\mu_{1}\rangle+(-1)\delta_{\mu_{1}<N}|N\rangle\right)\otimes \ldots \otimes \left(|\mu_{L}\rangle+(-1)\delta_{\mu_{L}<N}|N\rangle\right)
\end{equation}
therefore, projecting on $\langle \bm{\tau}|$ we have
\begin{equation}\label{Tensor-S1}
	\langle \bm{\tau}|	\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}|\bm{\mu}\rangle=\prod_{x=1}^{L}\left(\delta_{\mu_{x},\tau_{x}}+(-1)\delta_{\tau_{x},N}\sum_{i_{x}=1}^{N-1}\delta_{\mu_{x},i_{x}}\right)
\end{equation}
\begin{comment}
	Using the notation introduced in \eqref{notation1} and \eqref{notation2} we have that 
	\begin{equation}
		\begin{split}
			\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}&=\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L}}^{\tau_{1},\ldots,\tau_{L}}\right.
			\\&
			+\left.\sum_{b=1}^{L}(-1)^{b}\sum_{q_{1}=1}^{L}\sum_{q_{2}=q_{1}+1}^{L}\ldots\sum_{q_{b}=q_{b-1}+1}^{L}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{r}}=1}^{N-1}e_{\eta_{1},\ldots,\eta_{L}}^{\theta_{1},\ldots,\theta_{L}}\right\} 
		\end{split}
	\end{equation}
	where we have introduced, $\forall x\in \{1,\ldots,L\}$
	\begin{equation}
		%\eta_{x}=\begin{cases}
			%	\tau_{x}\qquad &\text{if}\quad x\neq 1,\ldots,b\\
			%	d_{x}\qquad &\text{otherwise}
			%\end{cases}\qquad
			\theta_{x}(b)=\begin{cases}
				\tau_{x}\qquad &\text{if}\quad x\notin \{q_{1},\ldots,q_{b}\}\\
				N \qquad &\text{otherwise}
			\end{cases}
		\end{equation}
		The dependence of $\theta(b)$ with respect to $b$ is usually understood for the sake of notation. 
		This is proved by induction. Indeed, for $L=1$ it is trivially true, by \eqref{S1-L1}. We assume that it is true for arbitrary $L\in \mathbb{N}$ and we prove that it holds for $L+1$. 
		\begin{align*}
			\left(\mathcal{S}_{1}^{-1}\right)^{\otimes L}\otimes \mathcal{S}_{1}^{-1}&=
			\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L}}^{\tau_{1},\ldots,\tau_{L}}\right. \\&+\left. \sum_{b=1}^{L}(-1)^{b}\sum_{q_{1}=1}^{L}\sum_{q_{2}=q_{1}+1}^{L}\ldots\sum_{q_{b}=q_{b-1}+1}^{L}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{r}}=1}^{N-1}e_{\theta_{1},\ldots,\theta_{L}}^{\eta_{1},\ldots,\eta_{L}}\right\}
			\\&\otimes
			\left\{ \sum_{\tau_{1}=1}^{N}e_{\tau_{L+1}}^{\tau_{L+1}}+(-1)\delta_{\tau_{L+1},N}\sum_{d_{L+1}=1}^{N-1}e_{1}^{d_{L+1}}\right\}
			\\&=
			\sum_{\tau_{1}=1}^{N}\ldots\sum_{\tau_{L+1}=1}^{N}\left\{e_{\tau_{1},\ldots,\tau_{L+1}}^{\tau_{1},\ldots,\tau_{L+1}}\right. \\&+\left. \sum_{b=1}^{L+1}(-1)^{b}\sum_{q_{1}=1}^{L+1}\sum_{q_{2}=q_{1}+1}^{L+1}\ldots\sum_{q_{b}=q_{b-1}+1}^{L+1}\left(\prod_{r=1}^{b}\delta_{s_{q_{r}},N}\right)\sum_{d_{q_{1}}=1}^{N-1}\ldots\sum_{d_{q_{b}}=1}^{N-1}e_{\theta_{1},\ldots,\theta_{L+1}}^{\eta_{1},\ldots,\eta_{L+1}}\right\}
		\end{align*}
		\begin{comment}
